---
title: "HR Medical Leave Intake"
format: html
server: shiny
page-layout: full
toc: false
---

```{r}
#| context: setup
#| include: false
required_packages <- c("shiny", "openxlsx")
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    paste0(
      "Please install the following package(s) before running or publishing: ",
      paste(missing_packages, collapse = ", ")
    )
  )
}

library(shiny)
library(openxlsx)

default_xlsx <- file.path("data", "medical_leave_records.xlsx")
excel_path <- path.expand(Sys.getenv("HR_MEDICAL_LEAVE_EXCEL", unset = default_xlsx))

allowed_leave_types <- c("Medical leave", "Annual leave")
headers <- c(
  "Submitted At",
  "Worker ID",
  "Name",
  "Leave type",
  "Leave date"
)

ensure_workbook <- function(path = excel_path) {
  dir.create(dirname(path), recursive = TRUE, showWarnings = FALSE)

  if (!file.exists(path)) {
    wb <- createWorkbook()
    addWorksheet(wb, "Records")
    writeData(
      wb,
      sheet = "Records",
      x = matrix(headers, nrow = 1),
      startRow = 1,
      colNames = FALSE
    )
    saveWorkbook(wb, path, overwrite = TRUE)
  }
}

append_row <- function(worker_id, name, leave_type, leave_date, path = excel_path) {
  ensure_workbook(path)

  wb <- loadWorkbook(path)
  sheet_name <- names(wb)[1]
  existing_rows <- nrow(readWorkbook(path, sheet = sheet_name, colNames = FALSE))
  next_row <- existing_rows + 1

  row_data <- data.frame(
    "Submitted At" = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    "Worker ID" = trimws(worker_id),
    "Name" = trimws(name),
    "Leave type" = trimws(leave_type),
    "Leave date" = format(as.Date(leave_date), "%Y-%m-%d"),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )

  writeData(
    wb,
    sheet = sheet_name,
    x = row_data,
    startRow = next_row,
    colNames = FALSE
  )

  saveWorkbook(wb, path, overwrite = TRUE)
}
```

::::: hero-grid
::: hero-copy
`Worker Self-Service`

# Medical Leave Submission

Workers can submit their details on this page. Each successful submission is written to the current Excel workbook path used by the app.

-   Prepare your `Worker ID`
-   Confirm your `Name`
-   Choose the correct `Leave type`
-   Select the exact `Leave date`
:::

::: hero-side
## Storage Note

When this app runs on `shinyapps.io`, it can write to an Excel file inside the app instance, but that local file is not guaranteed to persist across redeploys or instance changes.

Use a remote storage service if you need permanent HR records.
:::
:::::

::: deploy-note
For permanent cloud storage, replace local Excel writing with a remote backend such as Google Sheets, OneDrive/SharePoint, Dropbox, or a database.
:::

```{r}
div(
  class = "form-shell",
  div(
    class = "form-heading",
    h2("Leave Application Form"),
    p("Fill in the required fields and click Save To Excel.")
  ),
  div(
    class = "form-grid",
    div(class = "field-card", textInput("worker_id", "Worker ID")),
    div(class = "field-card", textInput("name", "Name")),
    div(
      class = "field-card",
      selectInput(
        "leave_type",
        "Leave type",
        choices = allowed_leave_types,
        selected = allowed_leave_types[1]
      )
    ),
    div(
      class = "field-card",
      dateInput(
        "leave_date",
        "Leave date",
        value = Sys.Date(),
        format = "yyyy-mm-dd"
      )
    )
  ),
  actionButton("submit", "Save To Excel", class = "primary-btn"),
  div(class = "status-wrap", uiOutput("status_ui")),
  div(
    class = "file-line",
    tags$strong("Excel file: "),
    code(normalizePath(excel_path, winslash = "/", mustWork = FALSE))
  )
)
```

```{r}
#| context: server
status_message <- reactiveVal(NULL)

observeEvent(input$submit, ignoreInit = TRUE, {
  current_session <- getDefaultReactiveDomain()
  worker_id <- trimws(input$worker_id)
  name <- trimws(input$name)
  leave_type <- trimws(input$leave_type)
  leave_date <- input$leave_date

  if (!nzchar(worker_id) || !nzchar(name) || !nzchar(leave_type) || is.null(leave_date)) {
    status_message(list(
      type = "error",
      text = "Please fill in Worker ID, Name, Leave type, and Leave date."
    ))
    return()
  }

  if (!leave_type %in% allowed_leave_types) {
    status_message(list(
      type = "error",
      text = "Please choose Medical leave or Annual leave."
    ))
    return()
  }

  parsed_date <- tryCatch(as.Date(leave_date), error = function(e) NA)
  if (is.na(parsed_date)) {
    status_message(list(
      type = "error",
      text = "Invalid date format."
    ))
    return()
  }

  tryCatch(
    {
      append_row(worker_id, name, leave_type, parsed_date)
      status_message(list(
        type = "success",
        text = "Record saved successfully."
      ))

      updateTextInput(current_session, "worker_id", value = "")
      updateTextInput(current_session, "name", value = "")
      updateSelectInput(current_session, "leave_type", selected = allowed_leave_types[1])
      updateDateInput(current_session, "leave_date", value = Sys.Date())
    },
    error = function(e) {
      status_message(list(
        type = "error",
        text = paste("Unable to save record:", conditionMessage(e))
      ))
    }
  )
})

output$status_ui <- renderUI({
  msg <- status_message()

  if (is.null(msg)) {
    return(NULL)
  }

  is_success <- identical(msg$type, "success")
  fg <- if (is_success) "#155724" else "#7a1f2a"
  bg <- if (is_success) "#dff6e5" else "#fde7ea"
  border <- if (is_success) "#b8e0c2" else "#f3bac1"

  div(
    class = "status-box",
    style = paste(
      sprintf("color: %s;", fg),
      sprintf("background-color: %s;", bg),
      sprintf("border: 1px solid %s;", border)
    ),
    msg$text
  )
})
```
